---
title: "EnviroLens Technical Report"
subtitle: "Air Pollution Environmental-Health Risk — Verdania MVP"
author: "EnviroLens Analytics"
date: today
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: true
  warning: false
---

# Purpose

This reproducible report documents the Verdania MVP methodology for integrating ambient air-quality measurements, facility respiratory aggregates, and community vulnerability data into the **AP-EHRI** risk index.

# Datasets

Three core datasets are integrated:

1. Environmental air-quality samples (PM2.5, NO₂)
2. Facility respiratory encounter rates
3. Population and socioeconomic indicators

See `metadata/data_dictionary/dictionary.md` and `metadata/data_inventory/catalogue.yaml`.

# Data quality

```{r}
#| label: dq
#| eval: false
source("../../analysis/r/analysis_helpers.R")
data <- load_exports("../../analysis/r/exports")
data$dq
```

When exports are available, overall DQ scores are summarized by dataset and dimension (completeness, validity, consistency, timeliness, uniqueness, geographic accuracy).

# Statistical analysis

Analyses include descriptive statistics of PM2.5 and respiratory rates, district comparisons, Pearson correlation between PM2.5 and respiratory encounter rates, and a simple linear regression.

```{r}
#| label: corr
#| eval: false
ct <- correlate_pm_resp(data$env, data$health)
print(ct)
fit <- simple_regression(data$env, data$health)
print(summary(fit))
```

# Risk index

AP-EHRI combines normalized PM2.5, respiratory rates, industrial proximity, vulnerability, poverty, access gaps, and reporting incompleteness. Weights and limitations are documented in `analysis/risk_model/README.md`.

# Limitations

- Synthetic geography and outcomes
- Respiratory encounters are not etiologically specific to air pollution
- Risk scores support prioritization, not clinical diagnosis

# Recommendations

1. Prioritize high / very-high AP-EHRI communities for targeted air monitoring
2. Investigate late-reporting facilities flagged by the DQ engine
3. Expand industrial emission inventories for proximity modelling
